Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-07-09 paper.Rmd
add para on model development and test set contamination

---
title: "Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*"
author: 
 - Andreas Beger^[Predictive Heuristics, adbeger@gmail.com.]
 - Richard K. Morgan^[V-Dem, rick.morgan2@gmail.com.]
 - Michael D. Ward^[Predictive Heuristics, Duke University, University of Washington, michael.don.ward@gmail.com. Corresponding author.]
date: "`r format(Sys.time(), '%d %B %Y')`"
thanks: John Ahlquist, Cassy L. Dorff, and Shahryar Minhas both provided helpful feedback on this project. All the code and several additional analyses can be found at our replication archive at https://github.com/andybega/Blair-Sambanis-replication.
abstract: We examine the research protocols in Blair and Sambanis (2020).  We find that there are several important mistakes and research decisions that determine their results. Fixing these mistakes results in a reversal of their claim that theory based models of escalation are better at predicting onsets of civil war than other kinds of models.  While their model is not very theoretical, it is outperformed by several of the ad hoc, putatively non-theoretical models they devise and examine.  
output: 
  bookdown::pdf_document2:
    keep_tex: true
    keep_md: true
    toc: false
#header-includes:
#  - \usepackage{setspace}\doublespacing
bibliography: "./references.bib"
nocite: |
  @cederman:weidmann:2017
---

@@

>>>>

---
title: "Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*"
author: 
 - Andreas Beger^[Predictive Heuristics, adbeger@gmail.com.]
 - Richard K. Morgan^[Varieties of Democracy Institute, University of Gothenburg, rick.morgan2@gmail.com.]
 - Michael D. Ward^[Predictive Heuristics, Duke University, University of Washington, michael.don.ward@gmail.com. Corresponding author.]
date: "`r format(Sys.time(), '%d %B %Y')`"
thanks: John Ahlquist, Cassy L. Dorff, and Shahryar Minhas both provided helpful feedback on this project. All the code and several additional analyses can be found at our replication archive at https://github.com/andybega/Blair-Sambanis-replication.
abstract: We examine the research protocols in Blair and Sambanis (2020).  We find that there are several important mistakes and research decisions that determine their results. Fixing these mistakes results in a reversal of their claim that theory based models of escalation are better at predicting onsets of civil war than other kinds of models.  While their model is not very theoretical, it is outperformed by several of the ad hoc, putatively non-theoretical models they devise and examine.  
output: 
  bookdown::pdf_document2:
    keep_tex: true
    keep_md: true
    toc: false
#header-includes:
#  - \usepackage{setspace}\doublespacing
bibliography: "./references.bib"
nocite: |
  @cederman:weidmann:2017
---

@@

Another important consideration when developing a predictive model is to do so without contaminating the out-of-sample test set. This occurs if test set performance is used to inform specification---predictive modeling's analogue to p-hacking---and special care is needed [e.g. see the approach by @colaresi:mahmood:2017]. B&S did pre-register their 2016-H1 forecasts, which verifies the integrity of the forecasts, but this does not assure the model development process. To be clear: there is no reason to suspect test set contamination in the escalation or other models. Rather, it is challenging to provide verifiable assurance thereof in a single analysis iteration that both develops and tests a model on data that was available contemporaneously.
